---
title: "Week 2: ggplot"
output:
  revealjs::revealjs_presentation:
    theme: white
    center: true
    transition: none
    incremental: true
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
library(readr)
knitr::opts_chunk$set(echo = TRUE)
```


## ggplot2 Overview
Why ggplot2?

### Advantages of ggplot2

>- consistent underlying grammar of graphics (Wilkinson, 2005)
>- very flexible
>- theme system for polishing plot appearance

## Grammar of Graphics
The basic idea: independently specify plot building blocks and combine them to create just about any kind of graphical display you want.

Building blocks of a graph include:

>- data
>- aesthetic mapping
>- geometric object
>- statistical transformations
>- faceting

## ggplot2 VS Base Graphics

Compared to base graphics, ggplot2

>- is more verbose for simple / canned graphics
>- is less verbose for complex / custom graphics
>- does not have methods (data should always be in a data.frame)
>- uses a different system for adding plot elements

## Aesthetic Mapping

Aesthetics are things that you can see. Examples include:

>- position (i.e., on the x and y axes)
>- color ("outside" color)
>- fill ("inside" color)
>- shape (of points)
>- linetype
>- size

Aesthetic mappings are set with the aes() function.

## Geometric Objects (geom)

Geometric objects are the actual marks we put on a plot. Examples include:

>- points (`geom_point`)
>- lines (`geom_line`)
>- boxplot (`geom_boxplot`)

A plot must have at least one geom; there is no upper limit. You can add a geom to a plot using the `+` operator

## Exercise 1: NCAA Basketball data

We will use data from the NCAA basketball tournament from 2011 - 2016 to create a polished figure.

```{r, message = F}
hoops <- read_csv('http://www.math.montana.edu/ahoegh/teaching/stat408/datasets/TourneyDetailedResults.csv')
hoops_2011 <- hoops %>% filter(Season >= 2011)
```

## NCAA Basketball Data Overview

```{r}
hoops_2011
```




## Graphical Primitives/ ggplot
```{r, fig.align='center', fig.height=2.75}
graph.a <- ggplot(data = hoops_2011, aes(Lfgm, Wfgm))
graph.a
```

## Adding Geoms: `geom_point()`
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_point()
```

## Adding Geoms: `geom_smooth()`
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_point() + 
  geom_smooth(method = 'loess', formula = 'y ~ x')
```

## Adding Geoms: `geom_rug()`
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_point() + 
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  geom_rug()
```

## Adding Geoms: `geom_density2d()`
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_point() + 
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  geom_rug() + geom_density2d()
```

## Adding Geoms: `geom_jitter()`
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_rug() + geom_density2d() + geom_jitter()
```

## Adding Geoms: `labs()`
```{r, fig.align='center', fig.height=2.65}
graph.a  + geom_rug() + geom_density2d() +
 geom_jitter() + 
  labs(x='Losing Team Field Goals Made', 
       y = 'Winning Team Field Goals Made')
```

## Scales: `xlim()` and `ylim()`
```{r, mysize=TRUE, fig.align='center',warning=FALSE, fig.height=2.6}
graph.a + geom_rug() + geom_density2d() +
 geom_jitter() + 
  labs(x='Losing Team Field Goals Made', 
       y = 'Winning Team Field Goals Made') +
  xlim(c(0,max(hoops_2011$Wfgm))) + ylim(c(0,max(hoops_2011$Wfgm)))
```

## Themes
There are a wide range of themes available in ggplot: [theme overview](https://ggplot2.tidyverse.org/reference/ggtheme.html)
```{r,fig.align='center', fig.height=2.75, echo = F}
graph.a + geom_point() + theme_light() + labs(x='Losing Team Field Goals Made', y = 'Winning Team Field Goals Made')
```

## More about aes
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_jitter(col = 'firebrick4')
```

## More about aes
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_jitter(aes(col = as.factor(Season)))
```


## More about aes
```{r, fig.align='center', fig.height=2.75}
graph.a + geom_jitter(aes(col = as.factor(Season)), size=3,alpha=.4)
```

## More about aes
```{r, echo = F, mysize=TRUE,fig.align='center'}
graph.a + geom_jitter(aes(shape = as.factor(Season),col=Wscore), size=3,alpha=.4)
```

## More about aes: Comment
```{r, eval = F, mysize=TRUE,fig.align='center'}
graph.a + 
  geom_jitter(aes(shape = as.factor(Season),col=Wscore),
              size=3,alpha=.4)
```

## Faceting
```{r, echo = F}
graph.a + geom_point() + facet_wrap(~Season)
```

## Faceting: Comment
```{r, eval = F}
graph.a + geom_point() + facet_wrap(~Season)
```

## Faceting
```{r, mysize=TRUE, eval=T, size='\\tiny',fig.align='center', fig.height=2.75}
graph.a + facet_wrap(~Season) + 
  geom_jitter(alpha=.5, aes(color=Wfgm3))
```

## Seattle Housing Data Set
Use the Seattle Housing Data Set [http://math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv](http://math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv) to create an interesting graphic, include informative titles, labels, and add an annotation.

```{r}
seattle_in <- read_csv('http://math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv')
```


## Exercise: ggplot2

Now use `ggplot2` to create an interesting graph using the Seattle Housing data set.

## Solution: ggplot2
```{r,fig.align='center',echo=F}
seattle_in$zipcode <- as.factor(seattle_in$zipcode)
graph.a <- ggplot(data = seattle_in, aes(sqft_living,price))
graph.a + geom_jitter(aes(col = zipcode))+ theme(plot.title = element_text(size=8), text = element_text(size=6)) + geom_smooth(method='loess')+ggtitle('Seattle Housing Sales: Price vs. Square Footage Living Space') + ylab('Sales Price (million dollars)') + xlab('Living Space (square foot)')+scale_y_continuous(breaks=c(seq(0,7000000,by=1000000)), labels=as.character(0:7)) +  annotate('text',3500,6000000, label = 'Housing price depends on zipcode', size=2) +annotate("rect", xmin = 0, xmax = 7250, ymin = 5500000, ymax = 6500000, alpha = .6) + geom_segment(aes(x=3500, xend=3500, y=5500000, yend=3000000),
                           arrow = arrow(length = unit(0.5, "cm")))
```

## Solution: ggplot2
```{r, eval = F}
seattle_in$zipcode <- as.factor(seattle_in$zipcode)
graph.a <- ggplot(data = seattle_in, aes(sqft_living,price))
graph.a + geom_jitter(aes(col = zipcode))+ 
  theme(plot.title = element_text(size=8), 
        text = element_text(size=6)) + geom_smooth(method='loess')+
  ggtitle('Seattle Housing Sales: Price vs. Square Footage Living Space') + 
  ylab('Sales Price (million dollars)') + 
  xlab('Living Space (square foot)')+
  scale_y_continuous(breaks=c(seq(0,7000000,by=1000000)),
                     labels=as.character(0:7)) +  
  annotate('text',3500,6000000, 
           label = 'Housing price depends on zipcode', size=2) +
  annotate("rect", xmin = 0, xmax = 7250, ymin = 5500000, ymax = 6500000, alpha = .6) + 
  geom_segment(aes(x=3500, xend=3500, y=5500000, yend=3000000),
                           arrow = arrow(length = unit(0.5, "cm")))
```
