---
title: "Lecture 2: Exercises"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```


## Reading Data files

```{r read.data}
Seattle <- read.csv(
  'http://math.montana.edu/ahoegh/teaching/stat408/datasets/SeattleHousing.csv', 
  stringsAsFactors = F)
```



## Exercise: Vectors

Create a vector with your first, middle, and last names.

## Exercise: Conditions in R
Note that `&` is an and operator

```{r, eval=FALSE}
pi > 3 & pi < 3.5
c(1,3,5,7) %in% 1:3
1:3 %in% c(1,3,5,7)
```



## Exercise: Subsetting
1. Create a new data frame that only includes houses worth more than $1,000,000.

2. From this new data frame what is the average living square footage of houses. 

## Exercise: Lists 
Explore the indexing with these commands.
```{r, mysize=TRUE, size='\\tiny', eval=F}
msu.info <- list( name = c('Waded Cruzado','Andy Hoegh'), 
         degree.from = c('University of Texas at Arlington','Virginia Tech'),
         job.title = c('President', 'Assistant Professor of Statistics'))
msu.info[1]
msu.info[[1]]
msu.info$name[2]
msu.info[1:2]
unlist(msu.info)
```

## Exercise: Arrays

Create an array of dimension 2 x 2 x 3, where each of the three 2 x 2 subarray (or matrix) is the Identity matrix.


## Exercise: Basic Plot

- Using only the subset of homes worth more than a million dollars, create a graphic.

